---
title: "**Appendix A**"
subtitle: "*Warming and enrichment interact to drive food-web simplification in freshwater systems*"
author: |
  | Willem Bonnaffé, Alain Danet, Camille Leclerc, Victor Frossard, 
  | Eric Edeline and Arnaud Sentis
date:  \today
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \usepackage{caption}
   - \DeclareCaptionLabelFormat{nospace}{#1#2}
   - \captionsetup[figure]{labelformat=nospace}
   - \captionsetup[table]{labelformat=nospace}
   - \usepackage[labelfont={bf}]{caption}
bibliography: "../doc/references_appendix.bib"
csl: "../doc/ecologyletters.csl"
label:
    fig: "**Figure S**"
    tab: "**Table S**"
---

```{r setup, include = FALSE}
library(tidyverse)
library(GGally)
library(ggspatial)
library(rgdal)
library(rnaturalearth)
library(sf)
library(kableExtra)
library(piecewiseSEM)

knitr::opts_chunk$set(echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  fig.pos = "H",  # pdf mode
  out.width = "80%"
  )
options(kableExtra.latex.load_packages = TRUE)


mypath <- rprojroot::find_package_root_file
source(mypath("R", "misc.R"))

myload(env_stream, env_lake, dataset_lake_stream, dir = mypath("data"))
dataset_lake_stream$type[dataset_lake_stream$type == 'stream'] <- 'Stream'
dataset_lake_stream$type[dataset_lake_stream$type == 'lake'] <- 'Lake'
```


\def\figurename{Figure S}
\def\tablename{Table S}

\newpage
# Community data
Fish communities in lakes were sampled using stratified random sampling design with multimesh gillnets [@Appelberg1995; @cen14757]. Sampling was conducted in late summer and early autumn, when most fishes are active but not spawning [@Prchalova2010]. The pre-defined number of nets per lake depended on area and depth of the lake [@Appelberg1995]. Specifically, benthic multimesh gillnets (12 panels of 5 to 55 mm knot-to-knot; 30 m long and 1.5 m high) were set in the benthic habitat, and for deep lakes (>10 m), sampling is completed using pelagic multimesh gillnets (11 panels of 6.5 to 55 mm knot-to-knot; 27.5 m long and 6 m high), which were lowered from the surface to 6 m each day. A total of 256 lakes were sampled at least once during the study period (Fig. S1).  
Fish communities in streams were sampled using a standardized electrofishing protocol based on stream size [@cen14011; @Irz2022]. Sampling was conducted from late spring through early autumn during the low flow period. Because the purpose of electrofishing is to obtain a sufficiently representative sample of the populations or communities present in the surveyed stations to allow comparative temporal and spatial analysis, we selected multi-species samplings of small streams and of deeper and/or wider streams. Small streams (depth <0.7m across the entire stream width) were sampled using anodes distributed evenly across the entire stream width, while larger streams were sampled along the bank either on foot or by boat (for more details see [@Danet2021; @Irz2022]). We selected sites for which protocol methods did not change across years, resulting in a total number of 373 stream sites sampled at least once during the study period (Fig. S1). 
\newline 
```{r scriptMapSamplingSites, include = FALSE , echo = FALSE, message = FALSE, warning = FALSE}

worldmap <- ne_countries(continent = 'europe', scale = 'large', type = 'countries', returnclass = 'sf')
fr <- data.frame(Country = "France", Focus = "YES")
world_joined <- left_join(worldmap, fr, by = c("name" = "Country"))
francemap <- ne_countries(country = 'france', scale = 'large', type = 'countries', returnclass = 'sf')
lakes <- readOGR(mypath("data-raw/ne_10m_lakes"), layer="ne_10m_lakes")
lakes <- st_as_sf(lakes)
rivers <- readOGR(mypath("data-raw/ne_10m_rivers_lake_centerlines"), layer="ne_10m_rivers_lake_centerlines")
rivers <- st_as_sf(rivers)
sf::sf_use_s2(FALSE)
francelakes <- st_intersection(st_as_sf(lakes), st_as_sf(francemap))
francerivers <- st_intersection(st_as_sf(rivers), st_as_sf(francemap))
BassinHydrographique <- readOGR(mypath("data-raw/BassinsHydrographiques_Métropole2019_BDTopage"), layer="BassinHydrographique")
BassinHydrographique <- st_as_sf(BassinHydrographique)
BassinHydrographique <- st_transform(BassinHydrographique, crs = st_crs(francemap))
BassinHydrographique <- st_intersection(st_as_sf(BassinHydrographique), st_as_sf(francemap))

```

```{r figMapSamplingSites, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 7, fig.align = 'center', fig.cap="Metropolitan France map showing the geographical distribution of the 629 sites, where taxonomic and environmental data have been collected. Green points represent sampled lakes (n = 256), purple points represent sampled streams (n = 373) and grey borders into France represent the 7 hydrographic basins."}

ggplot() +
  geom_sf(data = world_joined, fill = "white", color = "black", size = 0.05) +
  geom_sf(data = BassinHydrographique, fill = "#f0f0f0", color = "#bdbdbd", size = 0.25) +
  geom_sf(data = francelakes, col = '#6baed6', fill = '#6baed6', size = 0.05) +
  geom_sf(data = francemap, fill = alpha("white", 0), color = "black", size = 0.6) +
  annotation_scale(location = "bl", width_hint = 0.1, text_cex = 1.2) +
  annotation_north_arrow(which_north = "true", location = "tr", pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
  geom_point(data = dataset_lake_stream,
             aes(x = long, y = lat, fill = type), shape = 21, colour = "#000000", size = 3) +
  scale_fill_manual(values = c("#8da0cb", "#66c2a5"), name = "Communities samples") +
  coord_sf(xlim = c(-5, 9.75), ylim = c(41.3, 51.5), expand = FALSE) +
  theme(title = element_text(),
        legend.position = c(0.75, 0.08), legend.background = element_blank(), legend.key = element_blank(),
        legend.title = element_text(size = 15, face = "bold"),
        legend.text = element_text(size = 14),
        axis.text.x = element_blank(), axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),  axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.background = element_blank(),
        strip.background = element_rect(fill = "#000000", color = "#000000", size = 1, linetype = "solid"),
        panel.border = element_rect(colour = "#000000", fill = NA, size = 1))

```


\newpage
# Environmental variables
Organic enrichment often results from inorganic nutrient enrichment, which can lead to eutrophication [@Fitch2012; @OGorman2012]. We therefore examined the relationship between BOD and nutrients (i.e. total phosphorus and total nitrogen). Sampling and calculation of mean annual total phosphorus and total nitrogen followed the protocols and procedures of BOD for lakes and streams, respectively.  
Overall, significant positive correlations were found between BOD and total phosphorus or total nitrogen in both streams and lakes. In particular, we found a stronger relationship between BOD and total phosphorus in lakes (Pearson’s r = 0.852) than in streams (Pearson’s r = 0.355). As for the relationship between BOD and total nitrogen, the correlation was stronger in streams (Pearson’s r = 0.505) than in lakes (Pearson’s r = 0.337).

```{r figBODNutrients, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 7, fig.align = 'center', fig.cap="Scatterplot matrix of \\textit{BOD} and nutrients for the sampled lakes (in purple) and streams (in green). Scatterplots of each pair of numeric variable are drawn on the left part of the figure. Pearson correlation is displayed on the right. Variable distribution is available on the diagonal. The variable \\textit{TN} refers to total nitrogen and \\textit{TP} corresponds to total phosphorus."}

ggpairs(dataset_lake_stream, columns = c(5, 20:21), ggplot2::aes(colour = type),
        lower = list(continuous = wrap("smooth", alpha = 0.2)), upper = list(continuous = wrap("cor", size = 6)),
        columnLabels = c("BOD (mg/L)", "TN (mg/L)", "TP (mg/L)")) +
  scale_colour_manual(values = c("#8da0cb", "#66c2a5")) +
  scale_fill_manual(values = c("#8da0cb", "#66c2a5")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.position = "bottom",
        axis.text = element_text(size = 14, colour = "#000000"),
        strip.text.x = element_text(size = 16, face = "bold"),
        strip.text.y = element_text(size = 16, face = "bold"))

```

\newpage
# References
