---
title: "project_data_summary"
author: "Camille Leclerc"
date: "14/02/2022"
output: html_document
bibliography: "../doc/references.bib"
---

<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
      text-align: justify;
  }
td {  /* Table  */
  font-size: 18px;
}
h1.title {
  font-size: 38px;
  color: Black;
  text-align: justify;
}
h1 { /* Header 1 */
  font-size: 20px;
  color: #9e0142;
  text-align: justify;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: #d53e4f;
  text-align: justify;
}
code.r{ /* Code block */
    font-size: 18px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 18px;
}
</style>


```{r setup, include=FALSE}
mypath <- rprojroot::find_package_root_file
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, dev = "svg")
knitr::opts_knit$set(root.dir = mypath())
```


```{r, include = FALSE , echo=FALSE, message = FALSE, warning = FALSE}
## R sert-up
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggspatial)
library(kableExtra)
library(knitr)
library(magrittr)
library(PerformanceAnalytics)
library(rgdal)
library(rnaturalearth)
library(rstatix)
library(sf)
library(tidyverse)


mypath <- rprojroot::find_package_root_file
source(mypath("R", "misc.R"))


## Dataset
  # Lake data
  myload(yearly_avg_dbo5_lake, yearly_avg_temp_lake, dir = mypath("outputs"))    
  myload(network_lake_metrics, dir = mypath("data"))
  plando_spatial_coordinates <- read_csv("data-raw/plando_spatial_coordinates.csv")
  colnames(plando_spatial_coordinates)[1] <- "code_lac"
  lake_analysis <- read.csv("outputs/lake_list.txt", sep = "")
  lake_analysis$station_date <- paste(lake_analysis$code_lac, lake_analysis$camp_annee, sep = "_")
  
  
  dbo_lake <- yearly_avg_dbo5_lake
  dbo_lake$station_date <- paste(dbo_lake$code_lac, dbo_lake$year, sep = "_")
  yearly_avg_temp_lake$Lake_ID[yearly_avg_temp_lake$Lake_ID == "LER27"] <- "LER27a"
  temp_lake <- yearly_avg_temp_lake
  temp_lake$station_date <- paste(temp_lake$Lake_ID, temp_lake$year, sep = "_")
  
  env_lake <- left_join(lake_analysis, dbo_lake %>% dplyr::select(raw_value, station_date), by = "station_date")
  colnames(env_lake)[5] <- "dbo" 
  env_lake <- left_join(env_lake, temp_lake %>% dplyr::select(raw_value, station_date), by = "station_date")
  colnames(env_lake)[6] <- "temp"
  env_lake <- env_lake %>% drop_na(temp)
  sort(unique(env_lake$camp_annee))
  env_lake <- env_lake %>% dplyr:: filter(camp_annee < 2018)
  length(unique(env_lake$code_lac)) ; length(unique(env_lake$id_campagne))
  
  
  length(unique(env_lake$id_campagne)) ; length(unique(network_lake_metrics$id_campagne))
  network_lake_metrics$richness <- NA
  for(i in 1: nrow(network_lake_metrics)){
    data <- network_lake_metrics[[6]][[i]]
    data <- data %>% filter(species_name != "biof" & species_name != "det" & species_name != "macroph" & species_name != "phytob" & species_name != "phytopl" & species_name != "zoob" & species_name != "zoopl")
    data$species_name <- substr(data$species_name, 1, 3)
    network_lake_metrics$richness[i] <- length(unique(data$species_name))
    rm(data)
  }
  rm(i)
  metric_lake <- network_lake_metrics %>% dplyr::select(connectance, richness, nbnode, w_trph_lvl_avg, max_troph_lvl)
  
  
  data_lake <- left_join(env_lake, metric_lake, by = "id_campagne")
  data_lake <- left_join(data_lake, plando_spatial_coordinates, by = "code_lac")
  data_lake$type <- "lake"
  data_lake <- data_lake %>% dplyr::select(type, code_lac, id_campagne, camp_annee, dbo, temp, connectance, richness, nbnode, w_trph_lvl_avg, max_troph_lvl, lat_plando, long_plando)
  colnames(data_lake) <- c("type", "station", "opcod", "year", "dbo", "temp", "connectance", "richness", "nbnode", "w_trph_lvl_avg", "max_troph_lvl", "lat", "long")
  

  plando_spatial_coordinates <- unique(data_lake %>% dplyr::select(lat, long, type))


  rm(dbo_lake, yearly_avg_dbo5_lake, temp_lake, yearly_avg_temp_lake)
  
  
  # Stream data
  myload(op_analysis, station_analysis, dir = mypath("data"))
  myload(yearly_avg_water_chemical_stream, yearly_avg_water_temperature_stream, dir = mypath("data/environment"))    
  myload(network_stream_metrics, dir = mypath("data"))
  
  
  op_analysis <- op_analysis[which(op_analysis$date > "2004-12-31"),]
  op_analysis <- op_analysis[which(op_analysis$date < "2019-01-01"),]
  op_analysis$station_date <- paste(op_analysis$station, op_analysis$year, sep = "_")
  
  
  dbo_stream <- yearly_avg_water_chemical_stream %>% dplyr::filter(category == "DBO")
  dbo_stream$station_date <- paste(dbo_stream$id, dbo_stream$year, sep = "_")
  temp_stream <- yearly_avg_water_temperature_stream %>%
                  group_by(id, year) %>%
                  slice(which.min(value.predSE))
  temp_stream$station_date <- paste( temp_stream$id,  temp_stream$year, sep = "_")
  
  
  env_stream <- left_join(op_analysis %>% as.data.frame(.) %>% dplyr::select(opcod, station, year, station_date), dbo_stream %>% as.data.frame(.) %>% dplyr::select(press, station_date), by = "station_date")
  env_stream <- left_join(env_stream, temp_stream %>% as.data.frame(.) %>% select(value_corrected, station_date), by = "station_date")
  colnames(env_stream)[5] <- "dbo" ; colnames(env_stream)[6] <- "temp"
  env_stream <- env_stream %>% drop_na(dbo) %>% drop_na(temp)
  sort(unique(env_stream$year))
  length(unique(env_stream$station)) ; length(unique(env_stream$opcod))
  
  
  length(unique(env_stream$opcod)) ; length(unique(network_stream_metrics$opcod))
  network_stream_metrics <- network_stream_metrics %>% dplyr::filter(opcod %in% env_stream$opcod)
  length(unique(env_stream$opcod)) ; length(unique(network_stream_metrics$opcod))
  env_stream <- env_stream %>% dplyr::filter(opcod %in% network_stream_metrics$opcod)
  length(unique(env_stream$opcod)) ; length(unique(network_stream_metrics$opcod))
  
  network_stream_metrics$richness <- NA
  for(i in 1: nrow(network_stream_metrics)){
    data <- network_stream_metrics[[2]][[i]]
    network_stream_metrics$richness[i] <- length(unique(data$species))
    rm(data)
  }
  rm(i)
  metric_stream <- network_stream_metrics %>% dplyr::select(connectance, richness, nbnode, w_trph_lvl_avg, max_troph_lvl)

  
  stream_spatial_coordinates <- station_analysis %>%
    mutate(lat = unlist(map(station_analysis$geometry, 2)),
           long = unlist(map(station_analysis$geometry, 1)))
  stream_spatial_coordinates <- as.data.frame(stream_spatial_coordinates)
  colnames(stream_spatial_coordinates)[1] <- "station"
  
  
  data_stream <- left_join(env_stream, metric_stream, by = "opcod")
  data_stream <- left_join(data_stream, stream_spatial_coordinates, by = "station")
  data_stream <- as.data.frame(data_stream)
  data_stream$type <- "stream"
  data_stream <- data_stream %>% dplyr::select(type, station, opcod, year, dbo, temp, connectance, richness, nbnode, w_trph_lvl_avg, max_troph_lvl, lat, long)
  
  
  stream_spatial_coordinates <- unique(data_stream %>% dplyr::select(lat, long, type))
  
  
  data_final <- rbind(data_lake, data_stream)
  
  rm(dataset_lake_stream, dbo_stream, yearly_avg_water_chemical_stream, temp_stream, yearly_avg_water_temperature_stream, data_lake, data_stream, env_lake, env_stream, metric_lake, metric_stream)
```

&nbsp;

# **0. Hypothesis**

## **0.1 Short rational**

- Temperature and enrichment are predicted to have dramatic impacts on food-webs

- Stability is theoretically predicted to have an U-shape with temperature [@uszko_effects_2017], higher temperature leading to higher stability in "cold" food-webs, and to lower stability in "hot" food-webs

- Futhermore, metabolic theory predicts that higher trophic level suffer most
  from high temperature. Metabolic costs increase exponentially with temperature
  and higher trophic levels have generally higher body mass and then higher
  basal metabolic cost

- Enrichment (measured here by DBO), is predicted to decrease stability by
  increasing the amplitude of population fluctuations [@rosenzweig_paradox_1971] 
  
- Insterestingly, enrichment is predicted to dampens the effect of high
  temperature by increasing resources thereby decreasing starvation risk due to
  high temperature [@tabi_warming_2019]

- Stream and lakes are really different ecosystems:
  - Stream fishes are highly contrained by stream flow
  - Lakes are highly vertically stratified but offer limited opportunity for
    dispersal
  - Streams are horizontal stratified and offer opportunity for dispersal 

- Develop hypothesis:
  - about differences between lakes and streams... 
  - about the different network metrics

&nbsp;

## **0.2 Methods**

- Network metrics: Connectance, Max trophic level, Mean trophic level, number of nodes
- Environment: Temperature and DBO (to detail)

&nbsp;

## **0.3 Statistical ideas**

- Simple linear model: 

$N_i = \alpha + \beta_0F + \beta_1T_i  + \beta_2T_iF + \beta_3D_i + \beta_4D_iF + \beta_5D_iT_i + \beta_6D_iT_iF + \epsilon$

- With:
  - $i$: site i
  - $N_i$: Network metric of site $i$ 
  - $D$: DBO
  - $T$: Temperature
  - $F$: freshwater ecosystem type, lake or stream (logical variable ?)
  - $\alpha$: intercept 
  - $\beta$: slope
  - $\beta_2$, $\beta_4$, $\beta_5$: slope of simple interaction
    - $\beta_2$: link between network metrics and temperature change with
      ecosystem type ?
    - $\beta_4$: link between network metrics and DBO change with
      ecosystem type ?
    - $\beta_5$: link between network metrics and DBO change with
      temperature ? (hyp enrichment and temperature compensate each other)
  - $\beta_6$: slope of triple interaction; effect of temperature on the link
    between network metrics and DBO change with ecosystem type ?

- Structural Equation Model (one for each ecosystem type ?) [@gilbert_temperature_2019]:

```{r, echo=FALSE, message = FALSE, warning = FALSE}
DiagrammeR::grViz("digraph {
  graph [layout = dot, rankdir = TB]

  node [shape = rectangle]
  rec1 [label = 'Temperature']
  rec2 [label = 'DBO']
  rec3 [label =  'Species richness']
  rec4 [label = 'Number of nodes']
  rec5 [label = 'Mean trophic level']
  rec6 [label = 'Max trophic level']
  rec7 [label = 'Connectance']

  # edge definitions with the node IDs
  rec1 -> rec3
  rec1 -> rec4
  rec1 -> rec5
  rec1 -> rec6
  rec1 -> rec7

  rec2 -> rec3
  rec2 -> rec4
  rec2 -> rec5
  rec2 -> rec6
  rec2 -> rec7

  rec3 -> rec4
  rec3 -> rec5
  rec3 -> rec6
  rec3 -> rec7
  }",
  height = 500)
```
We certainly need to assess the collinearity among network descriptors in the
SEM, it may be high, but it was ok in [@danet_food-web_2021].

&nbsp;

# **1. Temporal and spatial coverage**
Due to biological and environmental data availability, the temporal coverage is from 2005 to 2017.  
256 sampled lakes (with a total of 401 obs) and 373 sampled stream stations (with a total of 2,791 obs) were considered.

```{r, include = FALSE , echo=FALSE, message = FALSE, warning = FALSE}
coordinates <- rbind(plando_spatial_coordinates, stream_spatial_coordinates)

  worldmap <- ne_countries(continent = 'europe', scale = 'large', type = 'countries', returnclass = 'sf')
  fr <- data.frame(Country = "France", Focus = "YES") 
  world_joined <- left_join(worldmap, fr, by = c("name" = "Country"))
  francemap <- ne_countries(country = 'france', scale = 'large', type = 'countries', returnclass = 'sf')
  lakes <- ne_download(scale = 10, type = 'lakes', category = 'physical')
  rivers <- ne_download(scale = 10, type = 'rivers_lake_centerlines', category = 'physical')
  sf::sf_use_s2(FALSE)
  francelakes <- st_intersection(st_as_sf(lakes), st_as_sf(francemap))
  francerivers <- st_intersection(st_as_sf(rivers), st_as_sf(francemap))
```
```{r, echo = FALSE, fig.width = 8, fig.height = 8}
  ggplot() +
    geom_sf(data = world_joined, fill = "white", color = "black", size = 0.05) +
    geom_sf(data = francemap, fill = gray(0.9), color = "black", size = 0.25) +
    geom_sf(data = francerivers, col = '#6baed6', size = 0.25) +  
    geom_sf(data = francelakes, col = '#6baed6', fill = '#6baed6', size = 0.05) +
    geom_point(data = coordinates,
             aes(x = long, y = lat, fill = type), shape = 21, colour = "#000000", size = 2) +
    scale_fill_manual(values=c("#482878FF", "#1F9E89FF"), name = "Communities samples") +
    coord_sf(xlim = c(-5, 9.75), ylim = c(41.3, 51.5), expand = FALSE) +
    theme(title = element_text(),
          plot.title = element_text(margin = margin(20,20,20,20), size = 18, hjust = 0.5),
          axis.text.x = element_blank(), axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.title.x = element_blank(),  axis.title.y = element_blank(),
          panel.grid.major = element_blank(), panel.background = element_blank(),
          strip.background = element_rect(fill = "#000000", color = "#000000", size = 1, linetype = "solid"),
          strip.text.x = element_text(size = 12, color = "#ffffff", face = "bold"),
          panel.border = element_rect(colour = "#000000", fill = NA, size = 1)) +
    theme(legend.position = c(0.16, 0.08), legend.background = element_blank(), legend.key = element_blank())

    
rm(plando_spatial_coordinates, stream_spatial_coordinates, coordinates, worldmap, fr, world_joined, francemap, lakes, rivers, francelakes, francerivers)
```

&nbsp;

# **2. Environmental variables**
```{r, echo=FALSE}
tbl.env <- data_final %>%
            dplyr::select(type, station, opcod, year, dbo, temp) %>%
            gather(key = 'env_variable', value = 'value', dbo:temp) %>%
            mutate(env_variable = replace(env_variable, env_variable == "temp", "temperature (°C)")) %>%
            mutate(env_variable = replace(env_variable, env_variable == "dbo", "dbo (mg/L)")) %>%
            select(type, value, env_variable) %>%
            filter_all(all_vars(!is.na(.))) %>%
            group_by(env_variable, type) %>%
            summarise(mean = sprintf("%.1f", mean(value, na.rm = TRUE)),
                      sd = sprintf("%.1f", sd(value, na.rm = TRUE)),
                      median = sprintf("%.1f", median(value, na.rm = TRUE)),
                      min = sprintf("%.1f", min(value, na.rm = TRUE)),
                      max = sprintf("%.1f", max(value, na.rm = TRUE))) %>%
            gather(key, value, mean:max) %>%
            unite(Group, type, key) %>%  
            mutate(Group = fct_relevel(Group, "lake_mean", "lake_sd", "lake_median", "lake_min", "lake_max",
                                       "stream_mean", "stream_sd", "stream_median", "stream_min", "stream_max")) %>%
            spread(Group, value)

tbl.env <- setNames(tbl.env, nm = sub(".+_", "", names(tbl.env)))

kable(tbl.env, align = c('l', rep('c', 10))) %>%
  kable_styling("striped") %>%
  add_header_above(c(" " = 1, "Lake" = 5, "Stream" = 5))   
```

&nbsp;

## **2.1. Correlation plots stream (left) and lake (right)**
```{r , echo = FALSE, fig.show = 'hold', fig.width = 4, fig.height = 4}
chart.Correlation(data_final %>%
                    dplyr::filter(type == "stream") %>%
                    dplyr::select(dbo, temp),
                  histogram = TRUE, method = "pearson")    

chart.Correlation(data_final %>%
                    dplyr::filter(type == "lake") %>%
                    dplyr::select(dbo, temp),
                  histogram = TRUE, method = "pearson") 
```

&nbsp;

# **3. Food web metrics**
```{r, echo=FALSE}
tbl.fw <- data_final %>%
            dplyr::select(type, station, opcod, year, connectance, richness, nbnode, w_trph_lvl_avg, max_troph_lvl) %>%
            gather(key = 'fw_metric', value = 'value', connectance:max_troph_lvl) %>%
            mutate(fw_metric = replace(fw_metric, fw_metric == "nbnode", "number of nodes")) %>%
            mutate(fw_metric = replace(fw_metric, fw_metric == "w_trph_lvl_avg", "mean trophic level")) %>%
            mutate(fw_metric = replace(fw_metric, fw_metric == "max_troph_lvl", "max trophic level")) %>%
            select(type, value, fw_metric) %>%
            filter_all(all_vars(!is.na(.))) %>%
            group_by(fw_metric, type) %>%
            summarise(mean = sprintf("%.1f", mean(value, na.rm = TRUE)),
                      sd = sprintf("%.1f", sd(value, na.rm = TRUE)),
                      median = sprintf("%.1f", median(value, na.rm = TRUE)),
                      min = sprintf("%.1f", min(value, na.rm = TRUE)),
                      max = sprintf("%.1f", max(value, na.rm = TRUE))) %>%
            gather(key, value, mean:max) %>%
            unite(Group, type, key) %>%  
            mutate(Group = fct_relevel(Group, "lake_mean", "lake_sd", "lake_median", "lake_min", "lake_max",
                                        "stream_mean", "stream_sd", "stream_median", "stream_min", "stream_max")) %>%
            spread(Group, value)

tbl.fw <- setNames(tbl.fw, nm = sub(".+_", "", names(tbl.fw)))

kable(tbl.fw, align = c('l', rep('c', 10))) %>%
  kable_styling("striped") %>%
  add_header_above(c(" " = 1, "Lake" = 5, "Stream" = 5)) 
```

&nbsp;

## **3.1. Correlation plots stream (left) and lake (right)**
```{r , echo = FALSE, fig.show = 'hold', fig.width = 4.5, fig.height = 4.5}
chart.Correlation(data_final %>%
                    dplyr::filter(type == "stream") %>%
                    dplyr::select(connectance, richness, nbnode, w_trph_lvl_avg, max_troph_lvl),
                  histogram = TRUE, method = "pearson")    

chart.Correlation(data_final %>%
                    dplyr::filter(type == "lake") %>%
                    dplyr::select(connectance, richness, nbnode, w_trph_lvl_avg, max_troph_lvl),
                  histogram = TRUE, method = "pearson") 
```

&nbsp;
&nbsp;

# **4. Environmental variables vs. Food web metrics**
```{r, echo=FALSE, fig.width = 10, fig.height = 15}
data <- data_final %>%
          dplyr::select(type, station, opcod, year, dbo, temp, connectance, richness, nbnode, w_trph_lvl_avg, max_troph_lvl) %>%
          gather(key = 'env_variable', value = 'env_value', dbo:temp) %>%
          mutate(env_variable = replace(env_variable, env_variable == "temp", "temperature (°C)")) %>%
          mutate(env_variable = replace(env_variable, env_variable == "dbo", "dbo (mg/L)")) %>%
          gather(key = 'fw_metric', value = 'metric_value', connectance:max_troph_lvl) %>%
          mutate(fw_metric = replace(fw_metric, fw_metric == "nbnode", "number of nodes")) %>%
          mutate(fw_metric = replace(fw_metric, fw_metric == "w_trph_lvl_avg", "mean trophic level")) %>%
          mutate(fw_metric = replace(fw_metric, fw_metric == "max_troph_lvl", "max trophic level"))

    
ggplot(data , aes(x = env_value, y = metric_value, fill = type, color = type)) +
  geom_point(shape = 21, colour = "#000000", size = 2, alpha = 0.2) +
  geom_smooth(method = "lm") +
  scale_fill_manual(values = c("#482878FF", "#1F9E89FF"), name = " ") +
  scale_colour_manual(values = c("#482878FF", "#1F9E89FF"), name = " ") +
  facet_grid(fw_metric ~ env_variable, scales = "free", switch = "y") +   # Put the y facet strips on the left
  scale_y_continuous(position = "right") +   # Put the y-axis labels on the right
theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.position = "bottom",
        axis.text = element_text(size = 16, colour = "#000000"), 
        axis.line.x = element_line(color = "#000000"), 
        axis.line.y = element_line(color = "#000000"),
        strip.text.x = element_text(size = 20, face = "bold"),
        strip.text.y = element_text(angle = 180, size = 20, face = "bold")) +
  labs(y = NULL, x = NULL)
```

# References {-}
